---
title: "NEW_IAEM_2"
output:
  pdf_document: default
  html_document: default
date: '2022-08-10'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
rm(list = ls())
```
```{r}
library(dplyr)
library(stringr)
library(readr)
library(tidyverse)
library(psych)
library(janitor)
library(tidyr)
library(rstanarm)
```
```{r}
setwd("/Users/calvinzhang/Desktop/IAEM project")

data2 <- read.csv("numeric.csv")
```
```{r}
new_data <- data2[-c(1:6,8:17)]
new_datafram <- data.frame(new_data)
```
```{r}
# filter people answered 1 or 2 and 4 or 5 
group1.index <- which(str_detect(new_datafram$Q2, "1|2")) 
group2.index <- which(str_detect(new_datafram$Q2, "4|5")) 
```
```{r}
# filter people answered 1 or 2, also contain 4 or 5. 
inter <- intersect(group1.index, group2.index) 
```
```{r}
# remove people answered 4 or 5 in people answered 1 or 2
group1.ind <- group1.index[-which( group1.index %in%  inter)] 
# remove people answered 1 or 2 in people answered 4 or 5
group2.ind <- group2.index[-which( group2.index %in%  inter)] 
# filter people only answered 3
which(new_datafram$Q2 =='3') 
```
```{r}
# create new columns with group #s
new_datafram$group <- NA
new_datafram$group[c(group1.ind, which(new_datafram$Q2 =='3'))] <- "group1"  
new_datafram$group[group2.ind] <- "group2" 
new_datafram$group[inter] <- "group3"
```
```{r}
# reorder the Q4-13 answers 
x <-c()
for(i in 5:14){
  x <- new_datafram[,i]
  ind1<- which(x=='1')
  ind2<- which(x=='5')
  ind3<- which(x=='2')
  ind4<- which(x=='4')
  x[ind1] <- '5'
  x[ind2] <- '1'
  x[ind3] <- '4'
  x[ind4] <- '2'
  new_datafram[,i] <- x
}
new_num <- na.omit(new_datafram)
```
```{r}
# create new column with sum 
numeric_sum <- apply(new_num[,5:14], 2, as.numeric)
sum <- c()
for (j in 1:nrow(numeric_sum)) {
  sum[j] = sum(numeric_sum[j,],na.rm = T)
}
new_num$sum <- sum
final_num <- new_num[-which(new_num$sum==0),]
df <- cbind(ID = 1:nrow(final_num), final_num)
df

```
```{r}
# histogram for number of people vs. sum of scores
h <- hist(df$sum,
     main = "Sum of question 4-13 scores",
     xlab = "Sum of scores",
     ylab = "Number of people",
     xlim = c(0,60),
     ylim = c(0,100),
     col = "darkmagenta", breaks = 12)
text(h$mids,h$counts,labels=h$counts, adj=c(0.5, -0.5))
```
```{r}
# details about dataset df (cleaned)
summary(df)
data_matrix <- data.matrix(df)
scale(data_matrix, center = T, scale = T)
```
#select group 1 and group 2 based on the sum using pivot wider
```{r}
df2 <- df %>%
  dplyr::select(ID,group,sum)
num_group <- df2 %>%
  pivot_wider(names_from = group, values_from = sum)
```
#details about dataset "num_group"
```{r}bb
summary(num_group)
scale(num_group, center = T, scale = T)
```
#check different groups means, median, range etc. 
```{r}
describe(num_group)

qd4 <- df %>%
  dplyr::select(ID,Q4_1,group)

qd4 <- qd4[!(is.na(qd$Q4_1) | qd$Q4_1==""), ]

qd4 <- as.data.frame(qd4)
qd4 <- qd4 %>%
   pivot_wider(names_from = group, values_from = Q4_1)
describe(qd4)

qd5 <- df %>%
  dplyr::select(ID,Q5_1,group)

qd5 <- qd5[!(is.na(qd5$Q5_1) | qd5$Q5_1==""), ]

qd5 <- as.data.frame(qd5)
qd5 <- qd5 %>%
   pivot_wider(names_from = group, values_from = Q5_1)
describe(qd5)

qd6 <- df %>%
  dplyr::select(ID,Q6_1,group)

qd6 <- qd6[!(is.na(qd6$Q6_1) | qd6$Q6_1==""), ]

qd6 <- as.data.frame(qd6)
qd6 <- qd6 %>%
   pivot_wider(names_from = group, values_from = Q6_1)
describe(qd6)

qd7 <- df %>%
  dplyr::select(ID,Q7_1,group)

qd7 <- qd7[!(is.na(qd7$Q7_1) | qd7$Q7_1==""), ]

qd7 <- as.data.frame(qd7)
qd7 <- qd7 %>%
   pivot_wider(names_from = group, values_from = Q7_1)
describe(qd7)

qd8 <- df %>%
  dplyr::select(ID,Q8_1,group)

qd8 <- qd8[!(is.na(qd8$Q8_1) | qd8$Q8_1==""), ]

qd8 <- as.data.frame(qd8)
qd8 <- qd8 %>%
   pivot_wider(names_from = group, values_from = Q8_1)
describe(qd8)

qd9 <- df %>%
  dplyr::select(ID,Q9_1,group)

qd9 <- qd9[!(is.na(qd9$Q9_1) | qd9$Q9_1==""), ]

qd9 <- as.data.frame(qd9)
qd9 <- qd9 %>%
   pivot_wider(names_from = group, values_from = Q9_1)
describe(qd9)

qd10 <- df %>%
  dplyr::select(ID,Q10_1,group)

qd10 <- qd10[!(is.na(qd10$Q10_1) | qd10$Q10_1==""), ]

qd10 <- as.data.frame(qd10)
qd10 <- qd10 %>%
   pivot_wider(names_from = group, values_from = Q10_1)
describe(qd10)

qd11 <- df %>%
  dplyr::select(ID,Q11_1,group)

qd11 <- qd11[!(is.na(qd11$Q11_1) | qd11$Q11_1==""), ]

qd11 <- as.data.frame(qd11)
qd11 <- qd11 %>%
   pivot_wider(names_from = group, values_from = Q11_1)
describe(qd11)

qd12 <- df %>%
  dplyr::select(ID,Q12_1,group)

qd12 <- qd12[!(is.na(qd12$Q12_1) | qd12$Q12_1==""), ]

qd12 <- as.data.frame(qd12)
qd12 <- qd12 %>%
   pivot_wider(names_from = group, values_from = Q12_1)
describe(qd12)

qd13 <- df %>%
  dplyr::select(ID,Q13_1,group)

qd13 <- qd13[!(is.na(qd13$Q13_1) | qd13$Q13_1==""), ]

qd13 <- as.data.frame(qd13)
qd13 <- qd13 %>%
   pivot_wider(names_from = group, values_from = Q13_1)
describe(qd13)
```
#Calculating alpha values
```{r}
df_alpha <- df %>%
  dplyr::select(ID, Q4_1, Q5_1, Q6_1, Q7_1, Q8_1, Q9_1, Q10_1, Q11_1, Q12_1, Q13_1)
df_alpha <- lapply(df_alpha,as.numeric)

df_alpha <- as.data.frame(df_alpha)
dim(df_alpha)

cro_alpha <- psych::alpha(df_alpha[,2:11], na.rm = TRUE, check.keys=TRUE)

cro_alpha
#kendalls taw: non-parametric correlation test
hist(df_alpha$Q8_1)
describe(num_group)
#mean scale for each question
#look at mean scales by age (are younger people or older people getting good scores), gender (are women or men getting good scores?) by race (white vs. nonwhite)???
describe(df_alpha)
```
```{r}
df3 <- lapply(df, as.numeric)
table(df3$Q16)
```
#recode Q16 race variable to white and non-white and compare group1 and group2 by race
```{r}
#0: non-white
#1: white
df4 <- df %>%
mutate(Q16 = recode(Q16, "1" = "0",
"2" = "0",
"3" = "0",
"4" = "0",
"5" = "1",
"6" = "0",
"7" = "NA"))

df4 <- df4 %>%
dplyr::select(ID,Q16,group,sum)
df4$Q16 <- as.factor(df4$Q16)
df4$Q16[df4$ID == 71] = "NA"

fit_race <- lm(sum ~ group + Q16, data = df4)
summary(fit_race)

mean_race_list <- tapply(df4$sum, df4$Q16, mean,
                         simplify = FALSE)
mean_race_list

#1 dependent variable: sum; #2 independent variables: group and race

#post1 <- stan_glm(sum ~ group + Q16, data = df4, 
                  #family = gaussian(link = "identity"), 
                  #seed = 12345)
#summary(post1)

#08.23.22 Data Services
df4 <- df4 %>%
  rowwise() %>%
  mutate(group_Q16 = paste(as.character(group), as.character(Q16)))

#using group1 1 (group 1 white as the reference)
#Linear Regression
df4$group_Q16 <- factor(df4$group_Q16,levels = c("group1 1","group1 0", "group2 0", "group2 1","group3 1,
                                                 group3 0"))
lm_race <- lm(sum ~ group_Q16, data=df4)
summary(lm_race)

gdw <- df4 %>%
  group_by(group_Q16) %>%
  summarise(
    count = n(),
    mean = mean(sum),
    median = median(sum),
    sd = sd(sum),
    se = sd/sqrt(n()),
    CI_lower = mean - (1.96 * se),
    CI_upper = mean + (1.96 * se),
  )
gdw
#Anova test
aov_race <- aov(sum ~ group_Q16, data=df4)
summary(aov_race)
coefficients(aov_race)
#pvalue: 0.197 show that there is no significant difference among these four groups. 

#Creating graph 
ggplot(data = df4,
       aes(x = group_Q16, y = sum))+
  geom_boxplot()


race_group <- df %>%
  dplyr::select(ID,Q16,group,sum)
race_group <- as.data.frame(race_group)
race_group
table(race_group$Q16)
nonwhite_group.index <- which(str_detect(race_group$Q16, "1|2|3|6"))
white_group.index <- which(str_detect(race_group$Q16, "5"))

race_group <- na.omit(race_group)
race_group$race <- NA
race_group$race[white_group.index] <- "white"
race_group$race[nonwhite_group.index] <-"non-white"
race_group <- na.omit(race_group)
table(race_group$race)

group1 <- grep("group1",df$group)
group2 <- grep("group2",df$group)
group3 <- grep("group3",df$group)

group1_white <- intersect(group1,grep("white", race_group$race)) #16
group1_nonwhite <-intersect(group1,grep("non-white", race_group$race)) #2

group2_white <- intersect(group2,grep("white", race_group$race)) #89
group2_nonwhite <-intersect(group2,grep("non-white", race_group$race)) #8

group3_white <- intersect(group3,grep("white", race_group$race)) #30
group3_nonwhite <-intersect(group3,grep("non-white", race_group$race)) #4





```
#compare group1 and group2 by gender
```{r}
df5 <- df %>%
  dplyr::select(ID,Q15,group,sum)
df5$Q15 <- as.factor(df5$Q15)
df5$Q15[df5$ID == 71] = "5"
df5 <- filter(df5, ID != 13 & ID != 71 & ID != 47 & ID != 49 & ID != 88)
#aggregate(sum ~ Q15 + group, df5, mean, exclude = NULL)


mean_sex_list <- tapply(df5$sum, df5$Q15, mean,
                        simplify = FALSE)
df5 <- df5 %>%
rowwise() %>%
mutate(group_Q15 = paste(as.character(group), as.character(Q15)))


gd <- df5 %>%
  group_by(group_Q15) %>%
  summarise(
    count = n(),
    mean = mean(sum),
    median = median(sum),
    sd = sd(sum),
    se = sd/sqrt(n()),
    CI_lower = mean - (1.96 * se),
    CI_upper = mean + (1.96 * se),
  )
gd
df5 <- df5 %>%
  rowwise() %>%
  mutate(group_Q15 = paste(as.character(group), as.character(Q15)))

#Linear Regression, using group1 1 (group1 Male) as reference
lm_gender<- lm(sum ~ group_Q15, data=df5)
summary(lm_gender)

#Creating graph
ggplot(data = df5,
       aes(x = group_Q15, y = sum))+
  geom_boxplot()


# 95% CI contain null value, if it contains then it's significant.

sex_group <- df %>%
  dplyr::select(ID,Q15,group,sum)

Male.index <- which(str_detect(sex_group$Q15, "1"))
Female.index <- which(str_detect(sex_group$Q15, "2"))
All_other.index <- which(str_detect(sex_group$Q15,"3|4"))
sex_group$sex <- NA
sex_group$sex[Male.index] <- "male"
sex_group$sex[Female.index] <-"female"
sex_group$sex[All_other.index] <- "all_other"
sex_group[sex_group == ""] <- NA
sex_group <- na.omit(sex_group)
sex_group <- as.data.frame(sex_group)

fit_sex <- lm(sum ~ group+sex, data = sex_group)
summary(fit_sex)

ggplot(data = sex_group,
       aes(x = sex, y = sum))+
  geom_boxplot()
mean_sex_list <- tapply(sex_group$sum, sex_group$sex, mean,
                        simplify = FALSE)
mean_sex_list

sex_group %>%
  group_by(sex) %>%
  summarise(
    count = n(),
    mean = mean(sum),
    median = median(sum)
  )

group1_male <- intersect(group1,grep("male", sex_group$sex)) #16
group1_female <-intersect(group1,grep("female", sex_group$sex)) #3

group2_male <- intersect(group2,grep("male", sex_group$sex)) #88
group2_female <-intersect(group2,grep("female", sex_group$sex)) #39

group3_male <- intersect(group3,grep("male", sex_group$sex)) #30
group3_female <-intersect(group3,grep("female", sex_group$sex)) #8
```

#compare group1 and group2 by age
```{r}
df6 <- df %>%
    dplyr::select(ID, Q14, group, sum)
df6$Q14[df6$ID == 71] = "5"
df6$Q14 <- as.factor(df6$Q14)

df6 <- filter(df6, ID != 51 & ID != 71 & ID != 47 & ID != 49 & ID != 88)

df6 <- df6 %>%
  rowwise() %>%
  mutate(group_Q14 = paste(as.character(group), as.character(Q14)))

#Linear Regression, using group1 1 (group1 19-35) as reference 
fit_age <- lm(sum ~ group_Q14, data=df6)
summary(fit_age)

#Creating graph
ggplot(data = df6,
       aes(x = group_Q14, y = sum))+
  geom_boxplot()

gda <- df6 %>%
  group_by(group_Q14) %>%
  summarise(
    count = n(),
    mean = mean(sum),
    median = median(sum),
    sd = sd(sum),
    se = sd/sqrt(n()),
    CI_lower = mean - (1.96 * se),
    CI_upper = mean + (1.96 * se),
  )
gda


age_group <- df %>%
  dplyr::select(ID,Q14,group,sum)
colnames(age_group)[2] <-"Age"
age_group <- as.data.frame(age_group)
age_group[age_group == ""] <- NA
age_group <- na.omit(age_group)
table(age_group$group)

fit_age <- lm(sum~group+Age, data = age_group)
summary(fit_age)

ggplot(data = age_group,
       aes(x = group, y = sum))+
  geom_boxplot()

age_group %>%
  group_by(group) %>%
  summarise(
    count = n(),
    mean = mean(sum),
    median = median(sum)
  )


```
#Comparing group1 and group2 by Hispanic race
```{r}
#recode non-hispanic to 0 and hispanic to 1 
df7 <- df %>%
    dplyr::select(ID, Q17, group, sum)
df7$Q17[df7$ID == 71] = "NA"
df7$Q17[df7$Q17 == 3] = "NA"
df7$Q17[df7$Q17 == 1] = "1"
df7$Q17[df7$Q17 == 2] = "0"
df7$Q17 <- as.factor(df7$Q17)

df7 <- filter(df7, ID != 13  & ID != 47 & ID != 49 & ID != 71 & ID != 88)

df7 <- df7 %>%
  rowwise() %>%
  mutate(group_Q17 = paste(as.character(group), as.character(Q17)))

#Linear Linear Regression, using group1 1 (group 1 Hispanic as reference)
df7$group_Q17 <- factor(df7$group_Q17,levels = c("group1 1","group1 0", "group2 0", "group2 1"))
fit_hispanic <- lm(sum ~ group_Q17, data = df7)
summary(fit_hispanic)

mean_hispanic_list <- tapply(df7$sum, df7$Q17, mean,
                             simplify = FALSE)
mean_hispanic_list

gdh <- df7 %>%
  group_by(group_Q17) %>%
  summarise(
    count = n(),
    mean = mean(sum),
    median = median(sum),
    sd = sd(sum),
    se = sd/sqrt(n()),
    CI_lower = mean - (1.96 * se),
    CI_upper = mean + (1.96 * se),
  )
gdh

#Creating graph
ggplot(data = df7,
       aes(x = group_Q17, y = sum))+
  geom_boxplot()

hispanic_group <- df %>%
  dplyr::select(ID,Q17,group, sum)

Hispanic.index <- which(str_detect(hispanic_group$Q17, "1"))
Non_Hispanic.index <- which(str_detect(hispanic_group$Q17, "2"))


hispanic_group$hispanic <- NA
hispanic_group$hispanic[Male.index] <- "hispanic"
hispanic_group$hispanic[Female.index] <-"non_hispanic"
hispanic_group <- as.data.frame(hispanic_group)
hispanic_group[hispanic_group == ""] <- NA
hispanic_group <- na.omit(hispanic_group)
table(hispanic_group$hispanic)

fit_hispanic <- lm(sum~group+hispanic, data = hispanic_group)
summary(fit_hispanic)

ggplot(data = hispanic_group,
       aes(x = hispanic, y = sum))+
  geom_boxplot()

mean_hispanic_list <- tapply(hispanic_group$sum, hispanic_group$hispanic, mean,
                             simplify = FALSE)
mean_hispanic_list

hispanic_group %>%
  group_by(hispanic) %>%
  summarise(
    count = n(),
    mean = mean(sum),
    median = median(sum)
  )


group1_his <- intersect(group1,grep("hispanic", hispanic_group$hispanic)) #16
group1_nonhis <-intersect(group1,grep("non_hispanic", hispanic_group$hispanic)) #4

group2_his <- intersect(group2,grep("hispanic", hispanic_group$hispanic)) #88
group2_nonhis <-intersect(group2,grep("non_hispanic", hispanic_group$hispanic)) #36

group3_his <- intersect(group3,grep("hispanic", hispanic_group$hispanic)) #30
group3_nohis <-intersect(group3,grep("non_hispanic", hispanic_group$hispanic)) #10
```

#Comparing means between group 1 and group 2 using Wilcoxon Test
```{r}
#group 1 is normally distributed
#group 2 is not normally distributed
shapiro.test(num_group$group1)
shapiro.test(num_group$group2)
shapiro.test(num_group$group3)

shapiro.test(log(num_group$group2))
wilcoxtest <- wilcox.test(num_group$group1, log(num_group$group2), 
                          alternative = "two.sided", mu=0, paired= FALSE)
wilcoxtest
```

#Comparing means between group 1 and group 2 using t test
```{r}
ttest <- t.test(num_group$group1~num_group$group2, alternative = "two.sided", mu = 0, paired= FALSE)
ttest
```
```{r}




```



