---
title: "NEW_IAEM_2"
output:
  pdf_document: default
  html_document: default
date: '2022-08-10'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
rm(list = ls())
```
```{r}
library(dplyr)
library(stringr)
library(readr)
library(tidyverse)
library(psych)
library(janitor)
library(tidyr)
library(rstanarm)
```
```{r}
setwd("/Users/calvinzhang/Desktop/IAEM project")

data2 <- read.csv("numeric.csv")
```
```{r}
new_data <- data2[-c(1:6,8:17)]
new_datafram <- data.frame(new_data)
```
```{r}
# filter people answered 1 or 2 and 4 or 5 
group1.index <- which(str_detect(new_datafram$Q2, "1|2")) 
group2.index <- which(str_detect(new_datafram$Q2, "4|5")) 
```
```{r}
# filter people answered 1 or 2, also contain 4 or 5. 
inter <- intersect(group1.index, group2.index) 
```
```{r}
# remove people answered 4 or 5 in people answered 1 or 2
group1.ind <- group1.index[-which( group1.index %in%  inter)] 
# remove people answered 1 or 2 in people answered 4 or 5
group2.ind <- group2.index[-which( group2.index %in%  inter)] 
# filter people only answered 3
which(new_datafram$Q2 =='3') 
```
```{r}
# create new columns with group #s
new_datafram$group <- NA
new_datafram$group[c(group1.ind, which(new_datafram$Q2 =='3'))] <- "group1"  
new_datafram$group[group2.ind] <- "group2" 
new_datafram$group[inter] <- "group3"
```
```{r}
# reorder the Q4-13 answers 
x <-c()
for(i in 5:14){
  x <- new_datafram[,i]
  ind1<- which(x=='1')
  ind2<- which(x=='5')
  ind3<- which(x=='2')
  ind4<- which(x=='4')
  x[ind1] <- '5'
  x[ind2] <- '1'
  x[ind3] <- '4'
  x[ind4] <- '2'
  new_datafram[,i] <- x
}
new_num <- na.omit(new_datafram)
```
```{r}
# create new column with sum 
numeric_sum <- apply(new_num[,5:14], 2, as.numeric)
sum <- c()
for (j in 1:nrow(numeric_sum)) {
  sum[j] = sum(numeric_sum[j,],na.rm = T)
}
new_num$sum <- sum
final_num <- new_num[-which(new_num$sum==0),]
df <- cbind(ID = 1:nrow(final_num), final_num)
df$group <- gsub("group3", "group1", df$group)
df
write.csv(df,"final_numeric.csv")
```
```{r}
# histogram for number of people vs. sum of scores
h <- hist(df$sum,
     main = "Sum of question 4-13 scores",
     xlab = "Sum of scores",
     ylab = "Number of people",
     xlim = c(0,60),
     ylim = c(0,100),
     col = "darkmagenta", breaks = 12)
text(h$mids,h$counts,labels=h$counts, adj=c(0.5, -0.5))
```
```{r}
# details about dataset df (cleaned)
summary(df)
data_matrix <- data.matrix(df)
scale(data_matrix, center = T, scale = T)
```
#select group 1 and group 2 based on the sum using pivot wider
```{r}
df2 <- df %>%
  select(ID,group,sum)
num_group <- df2 %>%
  pivot_wider(names_from = group, values_from = sum)
```
#details about dataset "num_group"
```{r}
summary(num_group)
scale(num_group, center = T, scale = T)
```
#check different groups means, median, range etc. 
```{r}
describe(num_group)
```
#Calculating alpha values
```{r}
df_alpha <- df %>%
  dplyr::select(ID, Q4_1, Q5_1, Q6_1, Q7_1, Q8_1, Q9_1, Q10_1, Q11_1, Q12_1, Q13_1)
df_alpha <- lapply(df_alpha,as.numeric)

df_alpha <- as.data.frame(df_alpha)
dim(df_alpha)
cro_alpha <- psych::alpha(df_alpha[,2:11], na.rm = TRUE, check.keys=TRUE)

cro_alpha
#kendalls taw: non-parametric correlation test
hist(df_alpha$Q8_1)
describe(num_group)
#mean scale for each question
#look at mean scales by age (are younger people or older people getting good scores), gender (are women or men getting good scores?) by race (white vs. nonwhite)???
describe(new_num)
```
```{r}
df3 <- lapply(new_num, as.numeric)
table(df3$Q16)
```
#recode Q16 race variable to white and non-white and compare group1 and group2 by race
```{r}
#0: non-white
#1: white
df4 <- df %>%
    mutate(Q16 = recode(Q16, "1" = "0",
                             "2" = "0",
                             "3" = "0",
                             "4" = "0",
                             "5" = "1",
                             "6" = "0",
                             "7" = "NA"))
df4 <- df4 %>%
  select(ID,Q16,group,sum)
df4$Q16 <- as.factor(df4$Q16)
df4$Q16[df4$ID == 71] = "NA"


df4 <- filter(df4, ID != 13 & ID != 18 & ID != 47 & ID != 49 & ID != 56 & ID != 71 & ID != 88 & ID != 135)
fit_race <- lm(sum ~ group + Q16, data = df4)
summary(fit_race)

mean_race_list <- tapply(df4$sum, df4$Q16, mean,
                         simplify = FALSE)
mean_race_list

df4 %>%
  group_by(Q16) %>%
  summarise(
    count = n(),
    mean = mean(sum),
    median = median(sum)
  )

#1 dependent variable: sum; #2 independent variables: group and race

#post1 <- stan_glm(sum ~ group + Q16, data = df4, 
                  #family = gaussian(link = "identity"), 
                  #seed = 12345)
#summary(post1)

#08.23.22 Data Services
df4 <- df4 %>%
  rowwise() %>%
  mutate(group_Q16 = paste(as.character(group), as.character(Q16)))

#using group1 1 (group 1 white as the reference)
#Linear Regression
df4$group_Q16 <- factor(df4$group_Q16,levels = c("group1 1","group1 0", "group2 0", "group2 1"))
lm_race <- lm(sum ~ group_Q16, data=df4)
summary(lm_race)

#Anova test
aov_race <- aov(sum ~ group_Q16, data=df4)
summary(aov_race)
coefficients(aov_race)
#pvalue: 0.197 show that there is no significant difference among these four groups. 

#Creating graph 
ggplot(data = df4,
       aes(x = group_Q16, y = sum))+
  geom_boxplot()

```
#compare group1 and group2 by gender
```{r}
df5 <- df %>%
  select(ID,Q15,group,sum)
df5$Q15 <- as.factor(df5$Q15)
df5$Q15[df5$ID == 71] = "5"
df5 <- filter(df5, ID != 13 & ID != 71 & ID != 47 & ID != 49 & ID != 88)
#aggregate(sum ~ Q15 + group, df5, mean, exclude = NULL)


mean_sex_list <- tapply(df5$sum, df5$Q15, mean,
                        simplify = FALSE)
mean_sex_list

df5 %>%
  group_by(Q15) %>%
  summarise(
    count = n(),
    mean = mean(sum),
    median = median(sum)
  )

df5 <- df5 %>%
  rowwise() %>%
  mutate(group_Q15 = paste(as.character(group), as.character(Q15)))

#Linear Regression, using group1 1 (group1 Male) as reference
lm_gender<- lm(sum ~ group_Q15, data=df5)
summary(lm_gender)

#Creating graph
ggplot(data = df5,
       aes(x = group_Q15, y = sum))+
  geom_boxplot()
```

#compare group1 and group2 by age
```{r}
df6 <- df %>%
    select(ID, Q14, group, sum)
df6$Q14[df6$ID == 71] = "5"
df6$Q14 <- as.factor(df6$Q14)

df6 <- filter(df6, ID != 51 & ID != 71 & ID != 47 & ID != 49 & ID != 88)

df6 <- df6 %>%
  rowwise() %>%
  mutate(group_Q14 = paste(as.character(group), as.character(Q14)))

#Linear Regression, using group1 1 (group1 19-35) as reference 
fit_age <- lm(sum ~ group_Q14, data=df6)
summary(fit_age)

#Creating graph
ggplot(data = df6,
       aes(x = group_Q14, y = sum))+
  geom_boxplot()

df6 %>%
  group_by(group) %>%
  summarise(
    count = n(),
    mean = mean(sum),
    median = median(sum)
  )
```
#Comparing group1 and group2 by Hispanic race
```{r}
#recode non-hispanic to 0 and hispanic to 1 
df7 <- df %>%
    select(ID, Q17, group, sum)
df7$Q17[df7$ID == 71] = "NA"
df7$Q17[df7$Q17 == 3] = "NA"
df7$Q17[df7$Q17 == 1] = "1"
df7$Q17[df7$Q17 == 2] = "0"
df7$Q17 <- as.factor(df7$Q17)

df7 <- filter(df7, ID != 13  & ID != 47 & ID != 49 & ID != 71 & ID != 88)

df7 <- df7 %>%
  rowwise() %>%
  mutate(group_Q17 = paste(as.character(group), as.character(Q17)))

#Linear Linear Regression, using group1 1 (group 1 Hispanic as reference)
df7$group_Q17 <- factor(df7$group_Q17,levels = c("group1 1","group1 0", "group2 0", "group2 1"))
fit_hispanic <- lm(sum ~ group_Q17, data = df7)
summary(fit_hispanic)

mean_hispanic_list <- tapply(df7$sum, df7$Q17, mean,
                             simplify = FALSE)
mean_hispanic_list

df7 %>%
  group_by(Q17) %>%
  summarise(
    count = n(),
    mean = mean(sum),
    median = median(sum)
  )

#Creating graph
ggplot(data = df7,
       aes(x = group_Q17, y = sum))+
  geom_boxplot()
```

#Comparing means between group 1 and group 2 using Wilcoxon Test
```{r}
#group 1 is normally distributed
#group 2 is not normally distributed
shapiro.test(num_group$group1)
shapiro.test(num_group$group2)


shapiro.test(log(num_group$group2))
wilcoxtest <- wilcox.test(num_group$group1, num_group$group2, alternative = "two.sided", mu=0, paired= FALSE)
wilcoxtest
```

#Comparing means between group 1 and group 2 using t test
```{r}
ttest <- t.test(num_group$group1, num_group$group2, alternative = "two.sided", mu = 0, paired= FALSE)
ttest
```
